---
title: "Codeliste viz for SPC .stat"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load tools of the trade

We load some useful library to interact with data in XML format, and with SDMX in particular

```{r}
library(tidyverse)
library(xml2)
```

For SDMX objext we are going to use a dedicated library

And for the plotting we can use an R library that creates Upset plots

```{r}
library(rsdmx)
```

```{r}
library(ComplexUpset)
library(ggplot2)
```

# Getting the data

## Get all codelist links

The main end point for the production PDH.stat rest request is https://stats-sdmx-disseminate.pacificdata.org/rest
And the codelist are served at codelist/all/all/latest?detail=allstubs

```{r}
base_url <- "https://stats-sdmx-disseminate.pacificdata.org/rest/"
query_all_cls <- "codelist/all/all/latest?detail=allstubs"
```

We get an xml document of all links to  codelists.
Unfortunately rSDMX does not play well with that kind of SDMX message.
Hence, we need to wrangle it by hand.

```{r}
all_cls_xml <- base_url |>
    paste0(query_all_cls) |>
    read_xml() |>
    xml_child(2) |>
    xml_child(1) |>
    xml_children()
```

Once we have a good list of all codelists, we flatten that to a dataframe

```{r}
dfo <- tibble(
    cid = all_cls_xml |> xml_attr("id"),
    name = all_cls_xml |> xml_text(),
    curl = all_cls_xml |> xml_attr("structureURL")
    ) 
```

## Get codes from codelist links

The next step will take all the url and retrieve the codes.
Thanks to rSDMX it's a one-liner.

```{r}
dfo <- dfo |>
    mutate(cl = map(curl, \(url) url %>% readSDMX %>% as_tibble))
```

## Write table to file

To write the codes to a csv we need to first unnest the tibble we created above.

```{r}
dfo <- dfo |>
    unnest_wider(cl) |>
    unnest_longer(id:description.en, keep_empty = TRUE) |>
    relocate(description.en, .after = name.en)
```

And we are ready to write.

```{r}
write_csv(dfo, paste0("all_codelists_",today(),".csv"))
```

Count indicators

```{r}
n_indicators <- dfo |>
    filter(cid |> str_detect("INDICATOR")) |>
    count() |>
    transmute(indicator_count = n) |>
    mutate(date = today())
```

and write to file

```{r}
if(file.exists("indicator_count.csv")) {
    write_csv(n_indicators, "indicator_count.csv", append = TRUE)
} else {
    write_csv(n_indicators, "indicator_count.csv", append = FALSE)
}
```

```{r}
dfo |>
    filter(cid |> str_detect("INDICATOR")) |>
    group_by(cid) |>
    count(sort = TRUE) |>
    transmute(indicator_count = n) |>
    mutate(date = today())
```

# Visualize overlaps

We can do some bipartite graph magic to visualize the overlaps.
This is a nice trick I learnt from [Bob Rudis](https://rud.is/books/21-recipes/visualizing-intersecting-follower-sets-with-upsetr.html)
and I updated a bit.

## Prepare data for viz

We first produce de-duplicated lists of all codes and codelists ids

```{r}
cl_c <- dfo |> select(cid,id)
unique_codes <- unique(cl_c$id)
unique_codelists <- unique(cl_c$cid)
```

Then for each code, we produce a dummy indicator of whether they are contained in a codelist or not and turn that into a tibble
```{r}
binaries <- unique_codelists |>
  map_dfc(
    \(this_codelist)
    ifelse(unique_codes %in% filter(cl_c, cid == this_codelist)$id, 1, 0) |> as_tibble()
    );
names(binaries) <- unique_codelists
```

We also do the transpose operation, creating a tibble of which codelist contains a certain code.
```{r}
rev_binaries <- unique_codes |>
  map_dfc(
    \(this_code)
    ifelse(unique_codelists %in% filter(cl_c, id == this_code)$cid, 1, 0) |> as_tibble()
    );
names(rev_binaries) <- unique_codes
```

## Plot the intersection

Plotting is also a one liner, but it might need some fine tuning for the graphical parameters
```{r}
codelist_overlap <- upset(binaries,cids, min_degree = 2, width_ratio = 0.2, height_ratio = 1.4, n_intersections=30)
codelist_overlap
```

```{r}
upset(rev_binaries,unique_codes, width_ratio = 0.2, height_ratio = 1.4, n_intersections=2)
```


```{r}
png(file="codelist_overlap.png", width = 1200, height = 800, points = 60) # or other device
codelist_overlap
dev.off()
```

## Get most common codes

```{r}
count_codes <- dfo |> count(id) |> arrange(-n)
```
