---
title: "Codeliste viz for SPC .stat"
output: github_document
---

link to all codelists <- https://stats-sdmx-disseminate.pacificdata.org/rest/codelist/all/all/latest?detail=allstubs

one example is here <- https://stats-sdmx-disseminate.pacificdata.org/rest/codelist/SPC/CL_CROPS_INDICATORS/1.0

```{r}
base_url <- "https://stats-sdmx-disseminate.pacificdata.org/rest/"
all_cls <- "codelist/all/all/latest?detail=allstubs"
```

for the one example

```{r}
crops_indicators_cl <- "codelist/SPC/CL_CROPS_INDICATORS/1.0"
```

# load weaponry

```{r}
library(tidyverse)
library(xml2)
library(rsdmx)
```

## gathering one codelist

```{r}
get_codelist <- function(url) {
   url |>
    readSDMX() |>
    as_tibble()
}

base_url |>
    paste0(crops_indicators_cl) |>
    get_codelist()
```


we want to get all the links!

```{r}
get_all_codelists <- function(url) {
   url |>
    read_xml() |>
    xml_child(2) |>
    xml_child(1) |>
    xml_children()
}

all_codelists <- base_url |>
    paste0(all_cls) |>
    get_all_codelists()
```

```{r}
build_CL_table <- function(all_codelists_xml) {

   table <- tibble(
    Cid = all_codelists_xml  |>
        xml_attr("id"),
    Curl = all_codelists_xml |>
        xml_attr("structureURL"),
    Cname = all_codelists_xml  |>
        xml_text()
    ) |>
    mutate(Codes = map(Curl, get_codelist))  |>
    unnest_wider(Codes) |>
    unnest_longer(id:description.en, keep_empty = TRUE)

    return(table)
}

Codelists <- all_codelists |> build_CL_table()
```

Wrap everything up in a function

```{r}
base_url <- "https://stats-sdmx-disseminate.pacificdata.org/rest/"
all_cls <- "codelist/all/all/latest?detail=allstubs"

get_codelist_table <- function(url) {
   table <- url |>
   get_all_codelists() |>
   build_CL_table()

   return(table)
}
```

```{r}
Codelists |>
    filter(Cid == "CL_COM_OBS_STATUS")
```